---
title: "Homework 2"
author: "DYLAN FRANCISCO"
date: "`r format(Sys.time(), '%B %d, %Y')`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Instructions

**Exercises:**  1-3 (Pgs. 49-50); 2-4 (Pg. 51); 2,4 (Pg. 54); 1-4 (Pg. 58)

**Submission:** Submit via an electronic document on Sakai. Must be submitted as a HTML file generated in RStudio. All assigned problems are chosen according to the textbook *R for Data Science*. You do not need R code to answer every question. If you answer without using R code, delete the code chunk. If the question requires R code, make sure you display R code. If the question requires a figure, make sure you display a figure. A lot of the questions can be answered in written response, but require R code and/or figures for understanding and explaining. 

```{r, include=FALSE}
library(tidyverse)
library(ggplot2)
library(nycflights13)
library(dplyr)
```

# Chapter 3 (Pgs. 49-50)

##  Exercise 1
```{r}
flights1 <- filter(flights, arr_delay >= 120)
flights1
flights2 <- filter(flights, dest %in% c("IAH", "HOU"))
flights2
flights3 <- filter(flights, carrier %in% c("UA", "AA", "DL"))
flights3
flights4 <- filter(flights, month %in% c(7, 8, 9))
flights4
flights5 <- filter(flights, arr_delay >= 120, dep_delay == 0)
flights5
flights6 <- filter(flights, dep_delay >= 60, arr_delay <= 30)
flights6
flights7 <- filter(flights, (dep_time >= 0000 & dep_time <= 0600))
flights7
```

The first time I tried this I spent 20 minutes trying to put all of these into one statement and trying to get it to work, only to finally realize that they should all be separate, at least I hope. May my pain be your humor.

##  Exercise 2
```{r}
flights7 <- filter(flights, between(dep_time, 0, 600))
flights7
```

Between() can be used here to solve the answer to part G.

##  Exercise 3
```{r}
flightsna <- filter(flights, is.na(dep_time))
flightsna
```

This may represent flights that may have been canceled.

```{r}
flightsna1 <- filter(flights, is.na(arr_time), !is.na(dep_time))
flightsna1
```


This may represent flights that, unfortunately, crashed.

# Chapter 3 (Pg. 51)

##  Exercise 2
```{r}
flightsdel <- flights %>%
  arrange(desc(dep_delay)) %>%
  select(dep_delay, everything())
flightsdel
flightsearly <- flights %>%
  arrange(dep_delay) %>%
  select(dep_delay, everything())
flightsearly
```

##  Exercise 3
```{r}
flightsfast <- flights %>%
  arrange(air_time) %>%
  select(air_time, everything())
flightsfast 
```

##  Exercise 4
```{r}
flightstrav <- flights %>%
  arrange(desc(distance)) %>%
  select(distance, everything())
flightstrav
flightstrav1 <- flights %>%
  arrange(distance) %>%
  select(distance, everything())
flightstrav1
```

Flights 51N380HA and 51N384HA flew the longest distance. Flights 3833N13989 and 4193N14972 flew the shortest distance.

Also glad I attended the pipeline lecture before doing this because it made it much easier.

# Chapter 3 (Pg. 54)

##  Exercise 2
```{r}
select(flights, dep_time, arr_time, dep_time)
```

If you include the name of variable multiple times, it will only show the first time.

##  Exercise 4
```{r}
select(flights, contains("TIME"))
select(flights, contains(toupper("time")))
```

The results of using contains("TIME") did not really surprise me. It seems like the helper ignores case and displays anything with the word "time" int it, regardless of case. I tried to use toupper() to only search for uppercase "TIME" but could not get it to work.

# Chapter 3 (Pg. 58)

##  Exercise 1
```{r}
flights8 <- transmute(flights, dep_time,
                      hour = dep_time%%100,
                      minute = dep_time%%100,
                      min_since_midnight = (hour*60) + minute)
flights8

```

##  Exercise 2
```{r}
flights9 <- transmute(flights, air_time,
                    arr_minus_dep = arr_time - dep_time)
flights9

flights10 <- transmute(flights, dep_hour = dep_time%%100,
                       dep_minute = dep_time%%100,
                       dep_min_after = (dep_hour*60) + dep_minute,
                       arr_hour = arr_time%%100,
                       arr_minute = arr_time%%100,
                       arr_min_after = (arr_hour*60) + arr_minute,
                       air_time, arr_minus_dep = arr_min_after - dep_min_after)


                  
flights10
```

Before running the code, I would expect the numbers to be the same.

After running, I noticed they are not, because arr_time and dep_time are measured in HHMM and air_time is in minutes.I thought a solution would be to convert the arrival and depature times to minutes and then do the calculation, but that did not end up working, however, I think that is still the solution.

##  Exercise 3
```{r}
flights11 <- transmute(flights, dep_time,
                       sched_dep_time,
                       dep_delay)
flights11
```

I would expect that if you add the dep_delay to the sched_dep_time, you would get the dep_time, which is correct.

##  Exercise 4
```{r}
flights12 <- flights %>%
  mutate(delay_rank = min_rank(desc(arr_delay))) %>%
  arrange(delay_rank) %>%
  select(delay_rank, everything())
flights12
```

These are the 10 most delayed flights. If there was a tie, the way min_rank handles it is that the tied values are given the same rank. The next, non-tied value is given the rank based off of its index in the column. 
