---
title: "Homework 2"
author: "DYLAN FRANCISCO"
date: "`r format(Sys.time(), '%B %d, %Y')`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Instructions

**Exercises:**  1-3 (Pgs. 49-50); 2-4 (Pg. 51); 2,4 (Pg. 54); 1-4 (Pg. 58)

**Submission:** Submit via an electronic document on Sakai. Must be submitted as a HTML file generated in RStudio. All assigned problems are chosen according to the textbook *R for Data Science*. You do not need R code to answer every question. If you answer without using R code, delete the code chunk. If the question requires R code, make sure you display R code. If the question requires a figure, make sure you display a figure. A lot of the questions can be answered in written response, but require R code and/or figures for understanding and explaining. 

```{r, include=FALSE}
library(tidyverse)
library(ggplot2)
library(nycflights13)
library(dplyr)
```

# Chapter 3 (Pgs. 49-50)

##  Exercise 1
```{r}
flights1 <- filter(flights, arr_delay >= 120)
flights1
flights2 <- filter(flights, dest %in% c("IAH", "HOU"))
flights2
flights3 <- filter(flights, carrier %in% c("UA", "AA", "DL"))
flights3
flights4 <- filter(flights, month %in% c(7, 8, 9))
flights4
flights5 <- filter(flights, arr_delay >= 120, dep_delay == 0)
flights5
flights6 <- filter(flights, dep_delay >= 60, arr_delay <= 30)
flights6
flights7 <- filter(flights, (dep_time >= 0000 & dep_time <= 0600))
flights7
```

The first time I tried this I spent 20 minutes trying to put all of these into one statement and trying to get it to work, only to finally realize that they should all be separate, at least I hope. May my pain be your humor.

##  Exercise 2
```{r}
flights7 <- filter(flights, between(dep_time, 0, 600))
flights7
```

Between() can be used here to solve the answer to part G.

##  Exercise 3
```{r}
flightsna <- filter(flights, is.na(dep_time))
flightsna
```

This may represent flights that may have been canceled.

```{r}
flightsna1 <- filter(flights, is.na(arr_time), !is.na(dep_time))
flightsna1
```


This may represent flights that, unfortunately, crashed.

# Chapter 3 (Pg. 51)

##  Exercise 2
```{r}
flightsdel <- flights %>%
  arrange(desc(dep_delay)) %>%
  select(dep_delay, everything())
flightsdel
flightsearly <- flights %>%
  arrange(dep_delay) %>%
  select(dep_delay, everything())
flightsearly
```

##  Exercise 3
```{r}
flightsfast <- flights %>%
  arrange(air_time) %>%
  select(air_time, everything())
flightsfast 
```

##  Exercise 4
```{r}
flightstrav <- flights %>%
  arrange(desc(distance)) %>%
  select(distance, everything())
flightstrav
flightstrav1 <- flights %>%
  arrange(distance) %>%
  select(distance, everything())
flightstrav1
```

Flights 51N380HA and 51N384HA flew the longest distance. Flights 3833N13989 and 4193N14972 flew the shortest distance.

Also glad I attended the pipeline lecture before doing this because it made it much easier.

# Chapter 3 (Pg. 54)

##  Exercise 2
```{r}
select(flights, dep_time, arr_time, dep_time)
```

If you include the name of variable multiple times, it will only show the first time.

##  Exercise 4
```{r}
select(flights, contains("TIME"))
```

The results of using contains("TIME") did not really surprise me. It seems like the helper ignores case and displays anything with the word "time" int it, regardless of case. 

# Chapter 3 (Pg. 58)

##  Exercise 1
```{r}
#
```

##  Exercise 2
```{r}
#
```

##  Exercise 3
```{r}
#
```

##  Exercise 4
```{r}
#
```